---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  dpi = 250,
  fig.dim = c(8,4),
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# adc

<!-- badges: start -->
[![adc status badge](https://txwri.r-universe.dev/badges/adc)](https://txwri.r-universe.dev)


[![license](https://img.shields.io/badge/license-GPL (>= 3)-lightgrey.svg)](https://choosealicense.com/)
[![R-CMD-check](https://github.com/TxWRI/adc/workflows/R-CMD-check/badge.svg)](https://github.com/TxWRI/adc/actions)
[![Codecov test coverage](https://codecov.io/gh/TxWRI/adc/branch/main/graph/badge.svg)](https://app.codecov.io/gh/TxWRI/adc?branch=main)

<!-- badges: end -->

adc provides functions to calculate discharge-based metrics that are useful 
in water quality concentration and flux regression equations.

## Installation

Install adc from CRAN:

```r
install.packages('adc')
```

The development version is available on https://txwri.r-universe.dev/adc and can be installed with:

``` r
install.packages('adc', repos = c(txwri = 'https://txwri.r-universe.dev'))
```

## Example

Flow anomalies represent how different the current discharge period is (current day, current week, current month, etc.) from previous periods (previous week, previous month, period of record, etc.). 

```{r example}
library(adc)
## example code is lavaca and includes dates and mean daily flow

data(lavaca)

x <- fa(lavaca$Flow,
        dates = lavaca$Date,
        T_1 = "1 month",
        T_2 = "1 year",
        clean_up = TRUE,
        transform = "log10")

plot(lavaca$Date, x, type = "l", xlab = "Date", ylab = "Anomaly [unitless]")
```

The packages also includes functions to calculate an exponentially weighted discounted flow, base-flow, and rate of change for mean daily streamflow. Functions generally work well using the `dplyr::mutate()` function to facilitate tidy workflows.
