---
title: "Introduction"
author: "Michael Schramm"
date: "`r Sys.Date()`"
output:
  html_document:
    math_method: "mathjax"
    mathjax: "local"
bibliography: references.yaml
csl: https://raw.githubusercontent.com/TxWRI/csl/main/twri-technical-report.csl
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "ragg_png",
  dpi = 250,
  fig.dim = c(8,4)
)
```

# Basic Functions

The adc package provides a handful of functions for developing
flow-based covariates useful in regression-based pollutant load
modeling. The primary package functions include:

-   `bf_sep_lh()` - uses the Lyne and Hollick filter to separate
    baseflow component from the quickflow component
    [@lyne_stochastic_1979]. Baseflow/Total Flow provides a baseflow
    index that is an estimate of the contribution of slow flow to total
    streamflow. This index can be useful for capturing relationships of
    subsurface solutes and streamflow. The function implements the
    approach described by @ladson_standard_2013.

-   `fa()` - calculates the flow anomaly, or the difference in flow for
    the current time period (day, week, month, year, etc.) from a prior
    period (days, weeks, months, years, or complete flow record)
    [@vecchia_trends_2009]. Positive values indicate the current period
    is wetter than the prior period. @zhang_improving_2017 demonstrate the use
    of different time periods to account for long-term, mid-term, and short-term
    flow anomalies.

-   `rate_of_change()` - is the estimate of the current instantaneous
    rate of change in daily streamflow. This is essentially an estimate
    of the difference in streamflow at time $T$ and $T-1$ 
    [@zhang_improving_2017]. Positive values mean streamflow is increasing and 
    negative values mean streamflow is decreasing. This term can capture 
    hysteresis effects in dissolved solutes and hydrograph limbs. This function 
    estimates $\frac{d Q}{d t}$ on the raw data or it will fit a cubic 
    regression spline between date and discharge when the argument 
    `smooth = TRUE` is used and return the derivative of the fitted function.

-   `sdf()` - the smooth discounted flow is an exponentially weighted
    average of streamflow values [@kuhnert_quantifying_2012]. This
    captures the integrated effect of all past flows with higher weight
    assigned to more recent flows than those that occurred long ago.

# Usage
All functions are intended to work on vectors of streamflow and
Date/Time data. Each function will return a numeric vector the same
length as the input arguments. An example dataset is provided:

```{r}
library(adc)
data("lavaca")

## run the lyne and hollick baseflow seperation filter
baseflow <- bf_sep_lh(discharge = lavaca$Flow,
                      a = 0.975)
## calculate the baseflow index
bfi <- baseflow/lavaca$Flow

head(bfi)

## calculate flow anomalies

#long term flow anomaly described in Zhang & Ball (2017)
ltfa <- fa(lavaca$Flow,
           dates = lavaca$Date,
           T_1 = "1 year",
           T_2 = "period",
           clean_up = TRUE,
           transform = "log10")
head(ltfa)

## rate of change
dqdt <- rate_of_change(lavaca$Flow, lavaca$Date)
head(dqdt)

## smooth discounted flow
ma <- sdf(lavaca$Flow, delta=0.95)
head(ma)

```


# Tidy Workflow

These functions can also be used in a tidy workflow:

```{r fig.height=8, fig.width=6, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(patchwork)

## create new column variable
lavaca |> 
  mutate(baseflow = bf_sep_lh(Flow, a = 0.975),
         ltfa = fa(Flow, Date, T_1 = "1 year",
                   T_2 = "period", clean_up = TRUE,
                   transform = "log10"),
         dqdt = rate_of_change(Flow, Date),
         ma = sdf(Flow)) |> 
  filter(Date >= as.Date("2004-01-01") & Date <= as.Date("2005-12-31"))-> df


##plot

## we are going to create a function to do this
plot_columns <- function(data, column, x_axis) {
  ggplot(data, aes_string(x = x_axis, y = column)) +
    geom_line(color = "steelblue") +
    labs(x = "Date", y = column) +
    theme_minimal()
}

## apply the function to each variable
plots <- lapply(c("Flow", "baseflow", "ltfa", "dqdt", "ma"),
                plot_columns,
                data = df,
                x_axis = "Date")

## use patchwork to put these all together
plots[[1]] / (plots[[2]] + plots[[3]]) / (plots[[4]] + plots[[5]])
```

# References
